---
title: "Exploratory Report"
date: "`r Sys.Date()`"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      fig.path = 'figures/')
```

```{r deps, message = FALSE}
library(knitr)
library(here)
library(tidyverse)
theme_set(theme_bw())
```

Machine learning algorithms used include: `r snakemake@params[['ml_methods']]`.
Models were trained with `r snakemake@params[['nseeds']]`
different random partitions of the data into training and testing sets using
`r snakemake@params[['kfold']]`-fold cross validation.

## Model Performance

```{r perf_plot, out.width='50%'}
include_graphics(snakemake@input[['perf_plot']])
```

### ROC & PRC Curves

```{r roc_plots, out.width='80%'}
include_graphics(snakemake@input[['roc_plots']])
```

## Hyperparameter Performance

```{r hp-plot, out.width='50%'}
include_graphics(snakemake@input[['hp_plot']])
```

## Memory Usage & Runtime

```{r runtime-plot, out.width='50%'}
include_graphics(snakemake@input[['bench_plot']])
```

Each model training run was given `r snakemake@params[['ncores']]` cores
for parallelization.

## Feature Importance

```{r load-feat-imp, message=FALSE}
alpha_level <- 0.05
dat <- read_csv(here('results', 'feature-importance_results.csv')) %>%
    rename(feature = names)

nseeds <- dat %>% pull(seed) %>% unique() %>% length()
ngroups <- dat %>% pull(test_group) %>% unique() %>% length()

signif_feats <- dat %>%
    filter(pvalue < alpha_level) %>%
    group_by(feature) %>%
    summarize(frac_sig = n() / (nseeds * ngroups))

feats <- dat %>%
    group_by(feature) %>%
    summarise(med_diff = median(perf_metric_diff),
              mean_diff = mean(perf_metric_diff),
              sd_diff = sd(perf_metric_diff)) %>%
    inner_join(signif_feats, by = "feature") %>%
    arrange(-mean_diff)
```

```{r plot-feats-histograms, message = FALSE, out.width = '50%'}
feats %>%
    ggplot(aes(frac_sig)) +
    geom_histogram()
feats %>% 
    ggplot(aes(mean_diff)) +
    geom_histogram()
feats %>% 
    ggplot(aes(frac_sig, mean_diff)) +
    geom_point(alpha = 0.5)
```

### Performance differences

```{r feat-imp-plot, out.width='50%'}
include_graphics(snakemake@input[['feat_plot']])
```

