---
title: 'Debugging log10-transformation of relative abundance plot'
date: '`r Sys.Date()`'
output:
  html_document:
    keep_md: false
    theme: paper
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs", output_format = "all") })
---
```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      fig.path = here::here('figures', 'debug/'),
                      dpi = 300
                      )
```

```{r deps}
library(here)
source(here('code', 'plotting-functions.R'))
```

```{r data}
metadat <- readxl::read_excel(here("data", "raw", "ml_metadata.xlsx")) %>%
  rename(sample = group)
abs_abun_dat <- data.table::fread(here("data", "raw", "sample.final.shared")) %>%
    rename(sample = Group) %>%
    right_join(metadat %>% select(sample, pos_cdiff_d1))
abs_abun_dat$total_counts <- rowSums(abs_abun_dat %>%
                                         select(starts_with("Otu")))
rel_abun_dat <- abs_abun_dat %>%
    pivot_longer(starts_with("Otu"),
                 names_to = "otu",
                 values_to = "count") %>%
    mutate(rel_abun = count / total_counts,
           pos_cdiff_d1 = capwords(pos_cdiff_d1)) %>%
    select(sample, pos_cdiff_d1, otu, rel_abun)
smallest_non_zero <- rel_abun_dat %>%
    filter(rel_abun > 0) %>%
    slice_min(rel_abun, n = 1) %>%
    pull(rel_abun) %>% .[1]

feat_dat <- read_csv(here("results", "feature-importance_results.csv"))
tax_dat <- schtools::read_tax(here("data", "processed", "final.taxonomy.tsv"))
top_feats_rel_abun <- get_top_feats(feat_dat, tax_dat, alpha_level = 0.05) %>%
    select(otu, label) %>%
    left_join(rel_abun_dat, by = "otu") %>%
    mutate(rel_abun_c = rel_abun + smallest_non_zero / 10,
           rel_abun_1 = rel_abun + 1)
```

### Original relative abundance, no transformation

```{r relabun_orig_no-transform}
top_feats_rel_abun %>% plot_rel_abun(xcol = rel_abun)
```

### Original relative abundance, log$_{10}$-transformed

```{r relabun_orig_log10}
top_feats_rel_abun %>% plot_rel_abun(xcol = rel_abun) + scale_x_log10()
```

Removes 363 rows since `log10(0) = -Inf`, so we need to add a constant

### Relative abundance + 1, no transformation

```{r relabun_1_no-transform}
top_feats_rel_abun %>% plot_rel_abun(xcol = rel_abun_1)
```

### Relative abundance + 1, log$_{10}$-transformed
```{r relabun_1_log10}
top_feats_rel_abun %>% plot_rel_abun(xcol = rel_abun_1) + scale_x_log10()
```

...that looks the same as above, just +1. 
Is there a problem in the scaling function?
What if I manually log-10 transform it?

```{r relabun_1_log10-manual}
top_feats_rel_abun %>% plot_rel_abun(xcol = log10(rel_abun_1))
```

The x axis numbers are different but the scale is still linear!
So maybe adding 1 is the problem?

### Relative abundance + tiny constant, no transformation

```{r relabun_c_no-transform}
top_feats_rel_abun %>% plot_rel_abun(xcol = rel_abun_c)
```

### Relative abundance + tiny constant, log$_{10}$-transformed

```{r relabun_c_log10}
top_feats_rel_abun %>% plot_rel_abun(xcol = rel_abun_c) + scale_x_log10()
```

## What does log$_{10}$-transforming do to a fake dataset?

```{r fake-data}
constant <- 0.0001
dat <- tibble(rel_abun = seq.int(0, 1, by = 0.01))  %>% 
    mutate(rel_abun_log10 = log10(rel_abun),
           rel_abun_1_log10 = log10(rel_abun + 1),
           rel_abun_c_log10 = log10(rel_abun + constant),
           rank = row_number())
set1colors <- RColorBrewer::brewer.pal(9, 'Set1')
dat %>% 
    pivot_longer(c(starts_with('rel_abun')), 
                 names_to = 'rel_abun_scale', 
                 values_to = 'rel_abun_value') %>% 
    mutate(rel_abun_scale = case_when(rel_abun_scale == 'rel_abun_1_log10' ~ 'log10(rel_abun + 1)',
                                      rel_abun_scale == 'rel_abun_c_log10' ~ 'log10(rel_abun + C)',
                                      rel_abun_scale == 'rel_abun_log10' ~ 'log10(rel_abun)',
                                      TRUE ~ rel_abun_scale)) %>% 
    ggplot(aes(rel_abun_value, rank, color = rel_abun_scale)) +
    geom_line(aes(size = rel_abun_scale), alpha = 0.5) +
    scale_size_manual(values = c(rel_abun = 1.5,
                                `log10(rel_abun)` = 1.5,
                                `log10(rel_abun + C)` = 2.5,
                                `log10(rel_abun + 1)` = 1.5)) +
    scale_color_manual(values=c(rel_abun = set1colors[9],
                                `log10(rel_abun)` = set1colors[5],
                                `log10(rel_abun + C)` = set1colors[2],
                                `log10(rel_abun + 1)` = set1colors[1])) + 
    theme_bw()
```

## Use log$_{10}$(rel_abun + C) and show the limit of detection

```{r relabun_c_log10_lod}
top_feats_rel_abun %>% 
    plot_rel_abun(xcol = rel_abun_c) + 
    geom_vline(xintercept = smallest_non_zero, linetype = 'dashed') +
    scale_x_log10() +
    labs(x = expression('Relative Abundance ('*log[10]+C*')'))
```

