---
date: '`r Sys.Date()`'
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
    includes:
      in_header: preamble.tex
  html_document:
    theme: paper
fontsize: 12pt
geometry: margin=1.0in
bibliography: references.bib
csl: msphere.csl
---
```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(message = FALSE,
                      fig.path = 'figures/',
                      dpi = 300
                      )
```

```{r deps}
library(here)
library(knitr)
library(tidyverse)
```

```{r data}
perf_dat <- read_csv(here('results','performance_results.csv'))
mean_auroc <- perf_dat %>% pull(AUC) %>% mean()
sd_auroc <- perf_dat %>% pull(AUC) %>% sd()
mean_auprc <- perf_dat %>% pull(prAUC) %>% mean()
sd_auprc <- perf_dat %>% pull(prAUC) %>% sd()
```

## Machine Learning Results

performance measured by the area under the receiver-operator characteristic curve (AUROC) and the area under the precision-recall curve (AUPRC).

Mean AUROC `r mean_auroc` (s.d. `r sd_auroc`)

Mean AUPRC `r mean_auprc` (s.d. `r sd_auprc`)

TODO feature importance

### Figure 5

TODO caption

## Machine Learning Methods

TODO describe pipeline [@topcuoglu_framework_2020]

mikropml version 1.2.1 [@topcuoglu_mikropml_2021]]

## References
